---
title: "Team Ken FJ! analysis"
author: "Monica Thieu"
date: "11/22/2018"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(googledrive)
require(magrittr)
require(tidyverse)
require(rvest)
# Apparently we need this shit now
options(mc.cores = parallel::detectCores())
```

Jeopardy!, the ultimate knowledge-based game show (in my opinion), lends itself very well to analysis about the game itself.

For the 2019 Jeopardy! All-Star Games, I started doing some rough analyses about what kinds of patterns appear in Jeopardy! games that we might be able to harness for our own use.

What this blog post **is:**

What this blog post **is not:**

## Setup

```{r}
# Load the data from me google drive
drive_download(as_id("https://docs.google.com/a/columbia.edu/spreadsheets/d/10l6IeI0VtWBaBtZDquGgIXUa0cSw-NnbbGW-F4hzfLE/edit"), path = "fj_teamken", type = "csv", overwrite = TRUE)

raw <- read_csv("fj_teamken.csv",
                col_names = c("timestamp",
                              "name",
                              "show_number",
                              "show_date",
                              "category",
                              "response",
                              "n_show_correct",
                              "comments",
                              "dd_sj_200",
                              "dd_sj_400",
                              "dd_sj_600",
                              "dd_sj_800",
                              "dd_sj_1000",
                              "dd_dj_400",
                              "dd_dj_800",
                              "dd_dj_1200",
                              "dd_dj_1600",
                              "dd_dj_2000"),
                skip = 1)

dds <- raw %>%
  filter(name == "Monica") %>%
  select(show_number, starts_with("dd")) %>%
  gather(key = "key", value = "clue_col", -show_number) %>%
  filter(!is.na(clue_col)) %>%
  separate(key, into = c("discard", "round", "clue_row")) %>%
  arrange(show_number, desc(round)) %>%
  group_by(show_number, round) %>%
  separate(clue_col, into = c("clue_col_1", "clue_col_2"), sep = "[,]", fill = "right") %>%
  select(-discard) %>%
  mutate(clue_row = as.integer(clue_row),
         clue_row = if_else(round == "sj",
                            clue_row / 200,
                            clue_row / 400),
         clue_row = as.integer(clue_row),
         clue_col_2 = str_sub(clue_col_2, start = 2L)) %>%
  gather(key = "discard", value = "clue_col", starts_with("clue_col")) %>%
  select(-discard) %>%
  filter(!is.na(clue_col)) %>%
  mutate(clue_col = as.integer(str_sub(clue_col, start = -1L)))

d <- raw %>%
  select(-timestamp, -show_date, -starts_with("dd")) %>%
  mutate(category = toupper(category),
         response = tolower(response),
         response = recode(response,
                           `incorrect, but thought of the correct answer and negged myself` = "self_neg",
                           `incorrect, but only because of spelling etc.` = "misspell",
                           `incorrect, but figured out what the clue was really asking for` = "almost"),
         response_binary = if_else(response == "correct", 1L, 0L),
         response_ternary = case_when(response == "correct" ~ 1,
                                      response %in% c("self_neg", "misspell") ~ 0.5,
                                      TRUE ~ 0),
         season = if_else(show_number >= 7586, 34, 26),
         category_ll = if_else(str_detect(comments, "ll \\("),
                               str_extract(comments, "\\(([^)]+)\\)"),
                               NA_character_),
         category_ll = str_sub(category_ll, start = 2L, end = -2L),
         category_ll = str_split(category_ll, pattern = ", "),
         category_ll_main = map_chr(category_ll, ~.x[1])) %>%
  group_by(season, show_number, category) %>%
  # Hack to get Monica's stuff on top
  arrange(desc(name)) %>%
  nest() %>%
  mutate(category_ll = map(data, ~.x$category_ll[[1]]),
         category_ll_main = map_chr(data, ~.x$category_ll_main[1]),
         data = map(data, ~.x %>% select(-starts_with("category_ll"))),
         toc = if_else(show_number %in% c(5916:5925, 7626:7635), TRUE, FALSE))

d_wide <- d %>%
  unnest(data, .preserve = "category_ll") %>%
  select(-comments, -starts_with("response_")) %>%
  mutate(name = recode(name,
                       Ken = "resp_kj",
                       Matt = "resp_mj",
                       Monica = "resp_mt")) %>%
  spread(name, response)

d_wide_complete <- d_wide %>%
  filter(!is.na(resp_kj), !is.na(resp_mj), !is.na(resp_mt))
```

## Person performance stuff

```{r}
d %>%
  filter(show_number %in% d_wide_complete$show_number, !is.na(category_ll_main), season == 34) %>%
  unnest(data) %>%
  ggplot(aes(x = name, fill = factor(response_binary))) +
  geom_bar() +
  scale_fill_manual(values = c("salmon", "springgreen2")) +
  facet_wrap(~category_ll_main, scales = "free_y") +
  labs(x = "Team member",
       y = "# of questions",
       fill = "Correct?") +
  theme_bw()
```

```{r}
d %>%
  filter(show_number %in% d_wide_complete$show_number, !is.na(category_ll_main), season == 34) %>%
  unnest(data) %>%
  ggplot(aes(x = name, fill = factor(response_ternary))) +
  geom_bar() +
  scale_fill_manual(values = c("salmon", "goldenrod1", "springgreen2")) +
  facet_wrap(~category_ll_main, scales = "free_y") +
  labs(x = "Team member",
       y = "# of questions",
       fill = "Correct?") +
  theme_bw()
```

```{r}
d %>%
  filter(toc) %>%
  unnest(data) %>%
  ggplot(aes(x = name, fill = factor(response_binary))) +
  geom_bar() +
  scale_fill_manual(values = c("salmon", "springgreen2")) +
  facet_wrap(~season, scales = "fixed") +
  labs(x = "Team member",
       y = "# of questions",
       fill = "Correct?") +
  theme_bw()
```

## Category frequency stuff

I realized that a big piece of information we could recover from the Final Jeopardy! clues that we all answered was _clue categories._

```{r}
d_wide_complete %>%
  filter(!is.na(category_ll_main)) %>%
  group_by(season) %>%
  count(category_ll_main) %>%
  mutate(n = if_else(season == 26, -1L * n, n),
         category_ll_main = fct_reorder(category_ll_main, n, .fun = max, .desc = TRUE)) %>%
  ggplot(aes(x = category_ll_main, y = n, fill = factor(season))) +
  geom_bar(stat = "identity") +
  # scale_fill_manual(values = c("salmon", "springgreen2")) +
  # facet_wrap(~category_ll_main, scales = "free_y") +
  coord_flip() +
  labs(x = "broad category",
       y = "# of questions",
       fill = "season",
       title = "FJ category frequencies by season") +
  theme_bw()
```

## Daily double stuff

```{r}
dds %>%
  group_by(round, clue_row, clue_col) %>%
  tally() %>%
  ungroup() %>%
  mutate(round = recode(round,
                        sj = "Single Jeopardy",
                        dj = "Double Jeopardy"),
         round = fct_rev(round)) %>%
  ggplot(aes(x = clue_col, y = clue_row, fill = n)) +
  geom_raster() +
  scale_y_reverse() +
  viridis::scale_fill_viridis() +
  facet_wrap(~ round) +
  labs(x = "Clue column",
       y = "Clue row",
       title = "Heatmap of DD incidence, most of S34",
       subtitle = "SJ slightly dimmer than DJ bc of more DDs left covered?",
       fill = "# of DDs") +
  theme_bw()
```

```{r}
dds %>%
  filter(round == "dj") %>%
  select(-clue_col) %>%
  group_by(show_number) %>%
  mutate(index = paste0("row_", 1:n())) %>%
  spread(index, clue_row) %>%
  mutate(row_diff = row_2 - row_1) %>%
  ggplot(aes(x = row_diff)) +
  geom_histogram(binwidth = 1) +
  labs(x = "ROW distance ($ amount) between DJ DDs") +
  theme_bw()
```


```{r}
dds %>%
  filter(round == "dj") %>%
  select(-clue_row) %>%
  group_by(show_number) %>%
  mutate(index = paste0("col_", 1:n())) %>%
  spread(index, clue_col) %>%
  mutate(col_diff = abs(col_2 - col_1)) %>%
  ggplot(aes(x = col_diff)) +
  geom_histogram(binwidth = 1) +
  labs(x = "COL distance (categories) between DJ DDs") +
  theme_bw()
```

```{r}
# If daily double locations were chosen completely independently of each other,
# what should the base rates of col-wise distance be? row-wise distance?

dd_sims <- tibble(id = 1:1000) %>%
  mutate(col_1 = sample(1:6, size = n(), replace = T),
         col_2 = map_int(col_1, ~sample((1:6)[1:6 != .x], size = 1, replace = T)),
         row_1 = sample(2:5, size = n(), replace = T),
         row_2 = sample(2:5, size = n(), replace = T),
         col_diff = abs(col_2 - col_1),
         row_diff = abs(row_2 - row_1))

dd_sims %>%
  ggplot(aes(x = row_diff)) +
  geom_histogram(binwidth = 1) +
  labs(x = "ROW distance ($ amount) between DJ DDs") +
  theme_bw()

dd_sims %>%
  ggplot(aes(x = col_diff)) +
  geom_histogram(binwidth = 1) +
  labs(x = "COL distance (categories) between DJ DDs") +
  theme_bw()
```


## extra lock 2 game shit

I was going to do a rough grid search across possible score sets to see the conformations that end in a game 2 lock, but the dataframe gets huge very quickly and stuff was running a little too slow for my taste.

```{r, eval = FALSE}
scores <- crossing(team1_game1 = 1:30,
                   team2_game1 = 1:30,
                   team3_game1 = 1:30) %>%
  filter(team1_game1 >= team2_game1, team1_game1 >= team3_game1)
```

## Daily Double extravaganza (as heard on Planet Money)

```{r big ass dd scrapey bitch, eval = FALSE}
seasons_raw <- 25:35 %>%
  paste0("http://www.j-archive.com/showseason.php?season=", .) %>%
  map(~read_html(.)) %>%
  map(~html_nodes(., "a"))

jscrape_raw <- seasons_raw %>%
  tibble(season = 25:35,
         raw = .) %>%
  mutate(text = map(raw, ~html_text(., trim = T)),
         link = map(raw, ~html_attr(., "href"))) %>%
  select(-raw) %>%
  unnest() %>%
  filter(grepl("showgame", link)) %>%
  mutate(web_id = as.integer(str_sub(link, start = -4L)),
         game_id = as.integer(str_sub(text, start = 2L, end = 5L)),
         these_htmls = map(web_id, ~read_html(paste0("http://j-archive.com/showgame.php?game_id=", .x))),
         these_nodes = map(these_htmls, ~html_nodes(.x, ".clue")))

jscrape <- jscrape_raw$these_nodes %>%
  map(~.x %>%
        html_text(trim = T) %>%
        tibble(raw = .) %>%
        # bye bye FJ, for now at least
        slice(-nrow(.))) %>%
  tibble(season = jscrape_raw$season,
         web_id = jscrape_raw$web_id,
         stuff = .) %>%
  filter(map_int(stuff, nrow) == 60) %>%
  unnest() %>%
  separate(raw, into = c("value", "order", "blank1", "blank2", "clue"), sep = "\n", fill = "right") %>%
  select(-starts_with("blank")) %>%
  mutate_all(str_trim, side = "both") %>%
  mutate(web_id = as.integer(web_id)) %>%
  # Splicing clue point value for the empty cells
  group_by(web_id) %>%
  mutate(idx = 1:n()) %>%
  filter(web_id != 3719) %>%
  mutate(column = rep(1:6, times = 10),
         row = rep(1:5, each = 6, times = 2),
         round = rep(c("sj", "dj"), each = 30),
         is_dd = if_else(startsWith(value, "DD"), TRUE, FALSE),
         nspaces_clue = str_count(clue, " ")) %>%
  select(-value, -idx)

write_csv(jscrape, "dd_scrape_25_35.csv")
```

```{r}
jscrape <- read_csv("dd_scrape_25_35.csv")
```

```{r}
jscrape %>%
  group_by(season, web_id, round, row, column) %>%
  summarize(n = sum(is_dd),
            mean_nspaces = mean(nspaces_clue, na.rm = TRUE)) %>%
  mutate(category_length = if_else(mean_nspaces > 10,
                                      "long",
                                      "short",
                                      missing = "short")) %>%
  group_by(round, row, column, category_length) %>%
  summarize(n = sum(n)) %>%
  ungroup() %>%
  mutate(round = recode(round,
                        sj = "Single Jeopardy",
                        dj = "Double Jeopardy"),
         round = fct_rev(round),
         n = na_if(n, 0),
         n = if_else(round == "Double Jeopardy",
                     n/2,
                     as.double(n))) %>%
  ggplot(aes(x = column, y = row, fill = n)) +
  geom_raster() +
  scale_y_reverse() +
  viridis::scale_fill_viridis() +
  # Remove category_length if not conditioning by this
  facet_grid(category_length ~ round) +
  labs(x = "Clue column",
       y = "Clue row",
       title = "Heatmap of DD incidence, most of seasons 25-35",
       subtitle = "DDs don't come up in wordplay categories",
       fill = "# of DDs") +
  theme_bw()
```

```{r}
jscrape %>%
  group_by(web_id, round, column) %>%
  summarize(mean_nspaces = mean(nspaces_clue, na.rm = TRUE),
            contains_dd = sum(is_dd)) %>%
  ungroup() %>%
  mutate(round = recode(round,
                        sj = "Single Jeopardy",
                        dj = "Double Jeopardy"),
         round = fct_rev(round)) %>%
  ggplot(aes(x = mean_nspaces, fill = factor(contains_dd))) +
  geom_density(alpha = 0.5) +
  labs(x = "# spaces per clue (proxy for word count)",
       title = "Category wordiness by DD incidence") +
  facet_grid(~round) +
  theme_bw()
```

```{r}
jscrape %>%
  filter(round == "dj", is_dd) %>%
  select(web_id, row) %>%
  group_by(web_id) %>%
  mutate(index = paste0("row_", 1:n())) %>%
  spread(index, row) %>%
  select(-row_3) %>%
  mutate(row_diff = row_2 - row_1) %>%
  ggplot(aes(x = row_diff)) +
  geom_histogram(binwidth = 1) +
  labs(x = "ROW distance ($ amount) between DJ DDs") +
  theme_bw()
```

```{r}
jscrape %>%
  filter(round == "dj", is_dd) %>%
  select(web_id, column) %>%
  group_by(web_id) %>%
  mutate(index = paste0("col_", 1:n())) %>%
  spread(index, column) %>%
  # This one game has 3 DDs listed but it must be wrong lol
  select(-col_3) %>%
  mutate(col_diff = abs(col_2 - col_1)) %>%
  ggplot(aes(x = col_diff)) +
  geom_histogram(binwidth = 1) +
  labs(x = "COL distance (categories) between DJ DDs") +
  theme_bw()
```

```{r}
dd_by_row <- jscrape %>%
  mutate(is_dd = as.integer(is_dd),
         # center at 10 words/clue (my wordplay cutoff)
         row = row - 3,
         row_quad = abs(row),
         column = column - 1,
         season = season - 35) %>%
  group_by(web_id, round, column) %>%
  mutate(category_length = if_else(mean(nspaces_clue) > 10,
                                      "long",
                                      "short",
                                      missing = "short")) %>%
  ungroup() %>%
  filter(!is.na(is_dd)) %>%
  nest(-round) %>%
  mutate(model = map(data, ~glm(is_dd ~ factor(row) + category_length + season - 1, 
                                family = binomial(),
                                data = .)))
```

```{r}
summary(dd_by_row$model[[2]])
```

```{r}
summary(dd_by_row$model[[1]])
```

```{r glm of prob of wordplay by column location}
category_length_by_col <- jscrape %>%
  group_by(season, web_id, round, column) %>%
  # signed so 1 is a short category
  summarize(category_length = mean(nspaces_clue, na.rm = T) <= 10) %>%
  ungroup() %>%
  filter(!is.na(category_length)) %>%
  mutate(season = season - 35) %>%
  nest(-round) %>%
  mutate(model = map(data, ~glm(category_length ~ factor(column) + season - 1, family = binomial(), data = .)))

category_length_by_col$model %>% map(summary)
```

Wordplay categories are less likely to be in the 1st column and more likely to be in the 6th column.

```{r}
dd_by_col <- jscrape %>%
  mutate(is_dd = as.integer(is_dd),
         # center at 10 words/clue (my wordplay cutoff)
         season = season - 35) %>%
  filter(!is.na(is_dd)) %>%
  group_by(season, web_id, round, column) %>%
  summarize(category_length = if_else(mean(nspaces_clue, na.rm = TRUE) > 10,
                                      "long",
                                      "short"),
            contains_dd = sum(is_dd)) %>%
  ungroup() %>%
  nest(-round) %>%
  mutate(model = map(data, ~glm(contains_dd ~ factor(column) + category_length + season, 
                                family = binomial(),
                                data = .)))
```

```{r}
summary(dd_by_col$model[[2]])
```

```{r}
summary(dd_by_col$model[[1]])
```

```{r}
dd_by_col %>%
  select(-model) %>%
  unnest() %>%
  filter(contains_dd == 1)

# This is category long-ness by column, categorically
jscrape %>%
  group_by(season, web_id, round, column) %>%
  # signed so 1 is a short category
  summarize(category_length = mean(nspaces_clue, na.rm = T) <= 10) %>%
  ungroup() %>%
  filter(!is.na(category_length)) %>%
  mutate(season = season - 35) %>%
  glm(category_length ~ factor(column) + season - 1, family = binomial(), data = .) %>%
  summary()
```

